---
title: "Analysis of Fishing Landings dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Potential dataset issues 

**Notes**:  
- Modified traps were only recorded in February.. is this a problem? The month of June for unmodified traps might be influencing some of these conclusions.. can we make that claim across all months?  
- Circle back to triple check the number of fish column in the QC script.    
- Make sure trap type names have been changed/edited as needed in QC script.    
- Will need a sanity check on reality of total catch values  

## Goal 

**1. Total catch per unit effort between modified and traditional traps. It would be great to see this as grams captured per trap set.**   

Results:    
- Modified traps had a significantly higher total catch per trap set. This pattern holds for all landing sites and individual fisherman (it does not appear that either of those variables influence this result).      
- Unmodified traps had a significantly higher weight in grams per trap set. This pattern holds for all individual fisherman, but Mayungu landing site might drive this result more than the other sites.      
- *insert statement about grams per trap vs catch per trap and trap type. finish stats below for this.* 

**2. Species catch per unit effort between modified and traditional traps. Take the top 3-5 species and run #1 for them separately.**

Results:  
- Top species caught across all surveys: *Siganus sutor* (38174), *Leptoscarus vaigiensis* (11739), *Lethrinus nebulosus* (6830), *Scarus ghobban* (5081), and *Siganus canaliculutus* (4135).   


3. Total mature fish catch per unit effort between modified and traditional traps. This will have to be for the top 3-5 species separately. Go to Fishbase and find the length at first maturity for that particular species, then assign each fish a "mature" or "immature" status in the data and calculate.    
4. Average length of catch versus length at first maturity (Lmat). Take the difference for each fish in the data against its length at first maturity and then calculate a weighted value for modified versus traditional traps where a value above 0 represents a fish above Lmat and a value below represents a fish below Lmat.   
5. Length frequency of top 3-5 species in modified versus traditional (different colors) with Lmat etc. indicators pulled from Fishbase.



## Contents 

- [**Reading in datafiles**](#data)  
- [**Total catch (grams) per unit effort (trap set)**](#catch_effort)  
- [**Calculate top species caught**](#species)  

## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
library(stats)
library(lme4)
library(car)
```

Read in the data frame that is the output of the QC script. 

```{r, message=FALSE, warning=FALSE}
# read in excel file
data <- read_excel("data/Fishlandings-cleaned-21052022-May.xlsx") #read in excel file 

# creating a month column based on operation date column 
data$month <- data$Operation_date #duplicating the operation date column to then modify in next steps 
data$month <- format(data$month, "%m") #extracting only the month in the new column
data$month <- as.numeric(data$month) #changing this column to numeric instead of a character (needed for next fxn)
data$month <- month.abb[data$month] #changing numeric months to month names 
```

## <a name="catch_effort"></a> **Total catch (grams) per unit effort (trap set)**

Grouping by fisher_id but this might be effective to group by enumerator once I have correct list of names. There are 3 boat trips recorded with the exact same fish data that are under 3 different fisher ID names but all the same enumerator.. come back to this in QC.

Goal: grams captured per trap set.

```{r}
modified_trap_df <- data %>% unite(survey_id, Operation_date, fisher_id, sep = " ", remove = FALSE) %>%
  dplyr::group_by(survey_id) %>% # group by survey id
  mutate(total_catch = sum(number_of_fish), #count the number of fish caught for each survey id
         grams_per_trap = total_weight_kg/total_traps_collected, #divide total weight for survey id by total traps 
         catch_per_trap = total_catch/total_traps_collected) %>% #divide total catch per survey id by total traps 
  dplyr::ungroup(survey_id) %>% #ungroup by this column  
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>% #subset for only modified and unmodified traps 
  select(survey_id, enumerator, trap_type, `No. of fishers in crew`, landing_site, total_catch, month, grams_per_trap, catch_per_trap) %>%
  distinct() 
```

### Plotting figures.

Catch per trap 

```{r}
# basic total catch per trap with no other variables 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Total catch per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by fisherman 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~enumerator) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by landing site
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~landing_site) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by month
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~month) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle
```

Grams per trap 

```{r}
# basic grams per trap plot with no other variables 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# visually seeing if this differs by fisherman 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~enumerator) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# visually seeing if this differs by landing site 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~landing_site) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by month
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~month) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle
```

The relationship between grams per trap and total catch per trap. 

```{r}
modified_trap_df %>% 
  ggplot(aes(x=catch_per_trap, y=grams_per_trap, color=trap_type)) +
  theme_bw() + xlab("total catch per trap") + ylab("grams per trap") +
  geom_smooth(aes(fill=trap_type), method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
  geom_point(aes(fill=trap_type), pch = 21, size=1)
```

### Statistics on the above. 

Default of t.test in R is a Welch t-test which is just an adaptation of t-test, and it is used when the two samples have possibly unequal variances. Use var.equal = TRUE or FALSE to specifiy the variances of the dataset. 

You can test equal variances with a Fisher's F-test. If p < 0.05 then we include var.equal = FALSE in below ttest. If p > 0.05 then we include var.equal = TRUE in below ttest.

We are using an unpaired two sample t-test for this dataset. 

```{r}
UN <- modified_trap_df %>% subset(trap_type == "MODIFIED") %>% na.omit()
MOD <- modified_trap_df %>% subset(trap_type == "UNMODIFIED") %>% na.omit()
```

#### Grams per trap 

```{r}
var.test(UN$grams_per_trap, MOD$grams_per_trap)
t.test(grams_per_trap~trap_type, data = modified_trap_df, var.equal = FALSE)
```

#### Total catch per trap  

```{r}
var.test(UN$catch_per_trap, MOD$catch_per_trap)
t.test(catch_per_trap~trap_type, data = modified_trap_df, var.equal = FALSE)
```

#### Total catch per trap vs weight in grams per trap. 

We use a linear mixed model for this so we can include other variables like fisherman and landing site. 

Grams per trap is log transformed.

I don't think this is the best way to do this.... circle back to a more proper model. Maybe just linear model for each type of trap? But then can't include name of fisherman and landing site as random factors.. 

```{r}
catch_model <- lmer(log(grams_per_trap) ~ catch_per_trap*trap_type + (1|enumerator) + (1|landing_site), na.action=na.omit, data=modified_trap_df)

qqPlot(residuals(catch_model)) 
hist(residuals(catch_model))

Anova(catch_model, ddf="lme4", type='III')
```


Left off at statistics and results for total catch per trap vs weight in grams per trap... question: does higher catch # result in higher weight in each modified and unmodified traps?

## <a name="species"></a> **Calculate top species caught**

Calculating which species were the most abundant across the entire survey.  

This might have to be number of fish caught per trap? So that the difference in # of traps for modified and unmodified is accounted for?

This is split for modified and unmodified so far.. but can be changed to combined.. the trend is about the same for most abundant type of fish in each trap.. 

```{r}
species_list <- data %>% select(scientific_name, number_of_fish, trap_type) %>% 
  na.omit() %>% 
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>% #subset for only modified and unmodified traps 
  dplyr::group_by(scientific_name, trap_type) %>%
  mutate(species_catch = sum(number_of_fish)) %>% select(-number_of_fish) %>% distinct()

# above 250 cut off includes top each count for each type of trap 
species_list %>% filter(species_catch > 250) %>%
  ggplot(., aes(x=scientific_name, y=species_catch, group = trap_type, color = trap_type)) + ylab("number of fish caught") + xlab("Genus species") +
  geom_point() + theme_bw() + theme(axis.text.x = element_text(angle = 60, hjust=1)) #Set the text angle

# print top 5 species from whole survey
species_list %>% group_by(trap_type) %>%                                    
  arrange(desc(species_catch))
```

Does this differ by type of trap?

```{r}

```



After this then subset for just top species and do the first goal with only those... 

