---
title: "Analysis of Fishing Landings dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal 

1. Total catch per unit effort between modified and traditional traps. It would be great to see this as grams captured per trap set.   
2. Species catch per unit effort between modified and traditional traps. Take the top 3-5 species and run #1 for them separately.    
3. Total mature fish catch per unit effort between modified and traditional traps. This will have to be for the top 3-5 species separately. Go to Fishbase and find the length at first maturity for that particular species, then assign each fish a "mature" or "immature" status in the data and calculate.    
4. Average length of catch versus length at first maturity (Lmat). Take the difference for each fish in the data against its length at first maturity and then calculate a weighted value for modified versus traditional traps where a value above 0 represents a fish above Lmat and a value below represents a fish below Lmat.   
5. Length frequency of top 3-5 species in modified versus traditional (different colors) with Lmat etc. indicators pulled from Fishbase.

## Contents 

- [**Reading in datafiles**](#data)  
- [**Total catch per unit effort**](#catch_effort)  

## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
```

Read in the data frame that is the output of the QC script. 

```{r, message=FALSE, warning=FALSE}
# read in excel file
data <- read_excel("data/Fishlandings-cleaned-21052022-May.xlsx") 
```

## <a name="catch_effort"></a> **Total catch per unit effort**

Grouping by fisher_id but this might be effective to group by enumerator once I have correct list of names. There are 3 boat trips recorded with the exact same fish data that are under 3 different fisher ID names but all the same enumerator.. come back to this in QC.

Goal: grams captured per trap set.

```{r}
data <- data %>% unite(survey_id, Operation_date, fisher_id, sep = " ") %>%
  dplyr::group_by(survey_id) %>% 
  mutate(total_catch = sum(number_of_fish),
         grams_per_trap = total_weight_kg/total_traps_collected)

modified_trap_df <- data %>% subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED")

modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + geom_boxplot() + theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1.1, hjust = 1.3)) #Set the text angle

t.test(grams_per_trap~trap_type, data = modified_trap_df)
```



