---
title: "Analysis of Fishing Landings dataset - alternate dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Prior Scripts 

QC: https://github.com/emmastrand/Kenya_SamakiSalama/blob/main/FishLandings/scripts/QC.md 


## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
library(stats)
library(lme4)
library(car)
library(forcats)
library(ggstatsplot)
library(ggridges)
library(ggbreak)
library(emmeans)
```

## Read in the data frame that is the output of the QC script. 

Summary information so far: 741 total surveys, 591 unmodified traps, 150 modified traps. 

```{r, message=FALSE, warning=FALSE}
# read in excel file
## 2,156 rows x 30 variables (this should match the final version of QC script)
data <- read.csv("alternate fishing/alternate_fishing.csv") %>% #read in excel file 
  dplyr::select(-X)

# creating new columns with month year and date
## survey ID is master identifier that is based on `date_collected_dd_mm_yyyy` column 
data <- data %>% 
  separate(date_collected_dd_mm_yyyy, c("year", "month", "day"), remove = FALSE) 

#changing this column to numeric instead of a character (needed for next fxn)
data$month <- as.numeric(data$month) 
#changing numeric months to month names
data$month <- month.abb[data$month] 

# changing levels of month (important for figures later on); only first three letter of month 
data$month <- factor(data$month, levels=c("Jan", "Feb","Mar","Apr","May",
                                          "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data <- data %>%
  mutate(household_group = case_when(
    BMU == "KANAMAI" ~ "Control",
    BMU == "KURUWITU" ~ "Social marketing",
    BMU == "TAKAUNGU" ~ "Social marketing"))
```


**Control households**   
- KANAMAI (141 surveys): November 2021 and January - May 2022
- Gear used: Spearguns
- Units of data: Per single fisherman / household

**Social marketing households**   
- KURUWITU (382 surveys): July - December 2021 and January - May 2022  
- TAKAUNGU (105 surveys): January - May 2022  
- Gear used: Both used spearguns and TAKAUNGU used monofilament 
- Units of data: Per single fisherman / household

### Read in the Group 3 

```{r}
##737 rows 
group3 <- read.csv("output/pertrap_final_df.csv", header=TRUE) %>% dplyr::select(-X) %>% 
  mutate(total_catch_per_fisherman = (CPUE*total_traps_collected)/`No..of.fishers.in.crew`,
        Total_Biomass_kg_per_fisherman = (calculated_biomass_pertrap*total_traps_collected)/`No..of.fishers.in.crew`,
        Total_value_KES_per_fisherman = (KSHpertrap*total_traps_collected)/`No..of.fishers.in.crew`,
        Calcium_mg_per_fisherman = (pertrap_Calcium*total_traps_collected)/`No..of.fishers.in.crew`,
        Iron_mg_per_fisherman = (pertrap_Iron*total_traps_collected)/`No..of.fishers.in.crew`,
        Omega3_g_per_fisherman = (pertrap_Omega3*total_traps_collected)/`No..of.fishers.in.crew`,
        Protein_g_per_fisherman = (pertrap_Protein*total_traps_collected)/`No..of.fishers.in.crew`,
        VitaminA_ug_per_fisherman = (pertrap_VitaminA*total_traps_collected)/`No..of.fishers.in.crew`,
        Selenium_ug_per_fisherman = (pertrap_Selenium*total_traps_collected)/`No..of.fishers.in.crew`,
        Zinc_mg_per_fisherman = (pertrap_Zinc*total_traps_collected)/`No..of.fishers.in.crew`)

##741 rows
length_data <- read.csv("output/length_persurvey.csv") %>% dplyr::select(survey_id, trap_type, median_length)

##737 rows -- should match the initial group3 dataframe 
group3 <- left_join(group3, length_data, by = c("survey_id", "trap_type"))

### will merge these data with group 1 and 2 at the end of this script 
```


## Check ranges and add filters 

63 unique species total:  
- 40 of those found in KANAMAI  
- 56 of those found in KURUWITU  
- 32 of those found in TAKAUNGU  

```{r}
hist(data$total_biomass_kg) ## 0.25 26.00
hist(data$take_home_weight_kg, na.rm = TRUE) ## 0-4
hist(data$total_value_KES) ## 40 - 4800 
hist(data$take_home_value_KES, na.rm = TRUE) ## 0-600 

range(data$`No..of.fishers.in.crew`) ## 1-3 people 

hist(data$number_of_fish) ## 1 - 19 fish 

#unique(data$scientific_name) ## 63 fish 
#data %>% subset(BMU == "TAKAUNGU") %>% dplyr::select(scientific_name) %>% distinct()
```

## Calculating median length 

```{r}
data <- data %>% 
  mutate(median_length = case_when(
    length_corrected == "0-10" ~ 5,
    length_corrected == "11-15" ~ 13,
    length_corrected == "16-20" ~ 18,
    length_corrected == "21-25" ~ 23,
    length_corrected == "26-30" ~ 28,
    length_corrected == "31-35" ~ 33,
    length_corrected == "36-40" ~ 38,
    length_corrected == "41-45" ~ 43,
    length_corrected == "46-50" ~ 48,
    length_corrected == "51-60" ~ 55.5,
    length_corrected == "61-70" ~ 65.5,
    length_corrected == "71-80" ~ 75.5,
    length_corrected == "81-90" ~ 85.5,
    length_corrected == ">90" ~ 100 ### circle back to what value to use here
  )) 

hist(data$median_length) ## majority fall within 10-40 cm 
```


## Loading Galligan and FishBase dataset 

Most species in the list above have metadata below. 

```{r}
fishbase_lifehistory <- read_excel("data/fishbase.xlsx", sheet = "life history") %>% #read in excel file 
  select(scientific_name, Lm, Lmax) %>% dplyr::rename(Lmat_fishbase = Lm) %>% dplyr::rename(Lmax_fishbase = Lmax) 

# rows that appear in species_list but not fishbase_lifehistory
## checking which fish I still need Fishbase information for (44 species)
# dplyr::setdiff(species_list, fishbase_lifehistory) %>% write.csv("data/add_fishbase_info.csv")

fishbase_biomass <- read_excel("data/fishbase.xlsx", sheet = "biomass") %>%
  dplyr::select(scientific_name, a, b)

fishbase <- full_join(fishbase_lifehistory, fishbase_biomass, by = "scientific_name")

galligan <- read.csv("data/SpeciesData_GatedTraps_Galligan_edited.csv", header=TRUE, sep = ",") %>%
  dplyr::rename(scientific_name = Species) #%>% dplyr::select(., scientific_name)
# dplyr::setdiff(species_list, galligan) 
### 11 fish found in our dataset that are not found in Galligan dataset 
## these won't be included in the nutrient or revenue analysis
## I think this is OK b/c all species have 113 or lower than 40 total catch

metadata <- full_join(fishbase, galligan, by = "scientific_name")

### PRE MERGE DATA = 2,156 X 34
### POST MERGE DATA = 2,156 X 82 
data2 <- left_join(data, metadata, by = "scientific_name")
```

## Calculating total biomass 

W=aL^b 

W = Biomass  
L = Length (in our case, median length)  
a = (taken from fishbase)
b = (taken from fishbase)

Read in data from fishbase. These units are in cm and grams. 

https://www.fishbase.de/manual/English/FishbaseThe_LENGTH_WEIGHT_table.htm

```{r}
### Switching to data3 which will filter out those fish with no nutrient or a and b values 
data2 <- data2 %>%  
  ## mutating a and b column to be numeric; this will create NAs b/c not all fish had a and b valueus in fishbase
  mutate_at(c('a', 'b'), as.numeric) %>%
  mutate(W_g = (a*(median_length^b))*number_of_fish,
         W_kg = W_g/1000) %>%
  group_by(survey_id) %>%
  filter(!is.na(W_kg)) %>%
  mutate(calculated_total_biomass = sum(W_kg)) %>% 
  ungroup()

range(data2$calculated_total_biomass) ## 0.001832111 kg - 18.201910245 kg
```

### Comparing reported vs. calculated 

```{r}
data2 %>% 
  select(survey_id, total_biomass_kg, calculated_total_biomass) %>%
  distinct() %>% na.omit() %>%
  mutate(comparison = if_else(total_biomass_kg > calculated_total_biomass, "REPORTED > CALC", "REPORTED < CALC"),
         difference = total_biomass_kg-calculated_total_biomass) %>%
  ggplot(., aes(x=comparison, y=difference)) + theme_bw() +
  geom_jitter(aes(color=comparison), size=1) + xlab("") + ylab("difference in value")
```

## Calculating take home kg/% 

```{r}
takehome <- data2 %>% dplyr::group_by(survey_id, destination) %>%   
  mutate(kg_perdestination = sum(W_kg)) %>% ungroup() %>% 
  dplyr::select(survey_id, BMU, household_group, destination, kg_perdestination) %>%
  distinct() %>% spread(destination, kg_perdestination) %>%
  mutate_all(., ~replace_na(.,0)) %>% dplyr::select(survey_id, BMU, household_group, HOME) %>%
  dplyr::rename(TakeHome_kg = HOME)
```

## Calculating trophic level 

```{r}
data2 <- data2 %>% group_by(survey_id) %>%
  mutate(mean.trophic = mean(TrophLevel, na.rm = TRUE))
```

## Calculating herbivore/carnivore presence 

```{r}
diet <- data2 %>% group_by(survey_id, Diet) %>%
  mutate(countper_diet = n()) %>% ungroup() %>%
  select(survey_id, BMU, household_group, Diet, countper_diet) %>% distinct() %>%
  group_by(survey_id) %>% spread(Diet, countper_diet) %>% dplyr::select(-`<NA>`) %>%
  mutate_all(., ~replace_na(.,0)) %>%
  mutate(total = sum(c(Carnivorous, `Herbivorous detritivorous`, `Macroalgal herbivorous`,
                        `Mobile Inverts`, `Omnivorous`, `Planktivorous`))) %>%
  mutate(Carnivorous = Carnivorous/total*100,
         `Herbivorous detritivorous` = `Herbivorous detritivorous`/total*100,
         `Macroalgal herbivorous` = `Macroalgal herbivorous`/total*100,
         `Mobile Inverts` = `Mobile Inverts`/total*100,
         `Omnivorous` = `Omnivorous`/total*100,
         `Planktivorous` = `Planktivorous`/total*100,
         Carnivorous_total = Carnivorous+`Mobile Inverts`) %>% 
  dplyr::select(-total, -Carnivorous, -`Mobile Inverts`) %>%
  dplyr::rename(Diet_percent_Herbivorous_Detritivorous = `Herbivorous detritivorous`) %>%
  dplyr::rename(Diet_percent_Herbivorous_Macroalgal = `Macroalgal herbivorous`) %>%
  dplyr::rename(Diet_percent_Omnivorous = Omnivorous) %>%
  dplyr::rename(Diet_percent_Planktivorous = Planktivorous) %>%
  dplyr::rename(Diet_percent_Carnivorous = Carnivorous_total)
  
head(diet)
```


## Calculating average length (cm)

```{r} 
data2 <- data2 %>% group_by(survey_id) %>%
  mutate(mean_length = mean(median_length))
```

## Calculating average value (KES)

```{r}
data2 <- data2 %>% 
  mutate(KSHperspp = Price_KSHPerkg*W_kg) %>%
  group_by(survey_id) %>%
  filter(!is.na(W_kg)) %>%
  mutate(calculated_total_value = sum(KSHperspp, na.rm=TRUE)) %>% 
  ungroup()
```

## Calculating average catch 

```{r}
data2 <- data2 %>%
  group_by(survey_id) %>%
  mutate(total_catch_per_fisherman = sum(number_of_fish))
```

## Calculating species richness

```{r}
data2 <- data2 %>%
  ## still grouped by survey id from above
  mutate(richness = n_distinct(scientific_name))
```

## Calculating nutrient information 

```{r}
data2 <- data2 %>%
  mutate(spp_Calcium = (W_g/100)*Calcium_mgPer100g,
         spp_Iron = (W_g/100)*Iron_mgPer100g,
         spp_Omega3 = (W_g/100)*Omega3_gPer100g,
         spp_Protein = (W_g/100)*Protein_gPer100g,
         spp_VitaminA = (W_g/100)*VitaminA_ugPer100g,
         spp_Selenium = (W_g/100)*Selenium_ugPer100g,
         spp_Zinc = (W_g/100)*Zinc_mgPer100g) %>%
  ### still grouped by survey id from above 
  mutate(Calcium_mg_per_fisherman = sum(spp_Calcium, na.rm = TRUE),
         Iron_mg_per_fisherman = sum(spp_Iron, na.rm = TRUE),
         Omega3_g_per_fisherman = sum(spp_Omega3, na.rm = TRUE),
         Protein_g_per_fisherman = sum(spp_Protein, na.rm = TRUE),
         VitaminA_ug_per_fisherman = sum(spp_VitaminA, na.rm = TRUE),
         Selenium_ug_per_fisherman = sum(spp_Selenium, na.rm = TRUE),
         Zinc_mg_per_fisherman = sum(spp_Zinc, na.rm = TRUE)) %>% ungroup()
```


## Summary and include Group 3 

```{r}
group_1_2 <- data2 %>% 
  dplyr::select(survey_id, year, month, day, BMU, household_group, calculated_total_biomass, mean.trophic,
                richness, total_catch_per_fisherman, calculated_total_value, mean_length, Calcium_mg_per_fisherman,
                Iron_mg_per_fisherman, Omega3_g_per_fisherman, Protein_g_per_fisherman, 
                VitaminA_ug_per_fisherman, Selenium_ug_per_fisherman,
                Zinc_mg_per_fisherman) %>%
  distinct() %>% 
  dplyr::rename(Total_Biomass_kg_per_fisherman = calculated_total_biomass) %>%
  dplyr::rename(Avg_Trophic_Level = mean.trophic) %>%
  dplyr::rename(Total_value_KES_per_fisherman = calculated_total_value) %>%
  dplyr::rename(Mean_length_cm = mean_length) %>%
  left_join(., takehome, c("survey_id", "BMU", "household_group")) %>%
  left_join(., diet, by = c("survey_id", "BMU", "household_group")) %>% ungroup() %>%
  ### subset to months of interest 
  subset(month == "Apr" | month == "May")

head(group_1_2)

### 161 surveys total for April and May combined 2022 
## Control = 41 surveys 
## Social marketing = 120 surveys 
```


### Subsetting group 3's data 

```{r}
diet_group3 <- read.csv("output/group3_diet.csv", header=TRUE) %>% dplyr::select(-X)
takehome_group3 <- read.csv("output/takehome_kg.csv", header=TRUE) %>% dplyr::select(-X) %>%
  dplyr::rename(TakeHome_kg = HOME)

## PRE FILTER = 737 surveys 
## POST FILTER 228 surveys
group3 <- group3 %>% 
  subset(month == "Apr" | month == "May") %>%
  subset(year == "2022") %>%
  dplyr::select(survey_id, BMU, trap_type, total_traps_collected, year, month, day, `No..of.fishers.in.crew`,
                richness, total_catch_per_fisherman, Total_Biomass_kg_per_fisherman, Total_value_KES_per_fisherman,
                Calcium_mg_per_fisherman, Iron_mg_per_fisherman, Omega3_g_per_fisherman, Protein_g_per_fisherman,
                VitaminA_ug_per_fisherman, Selenium_ug_per_fisherman, Zinc_mg_per_fisherman, mean.trophic, median_length) %>%
  mutate(household_group = case_when(
    trap_type == "Control" ~ "Social Marketing + Traps: Control",
    trap_type == "Experimental" ~ "Social Marketing + Traps: Experimental"
  )) %>% 
  dplyr::rename(No_fishermen = `No..of.fishers.in.crew`) %>%
  dplyr::rename(Mean_length_cm = median_length) %>%
  dplyr::rename(Avg_Trophic_Level = mean.trophic) 

group3 <- group3 %>% 
  left_join(., diet_group3, by = c("survey_id", "trap_type", "BMU")) %>%
  left_join(., takehome_group3, by = c("survey_id", "trap_type"))


group3 %>% subset(trap_type == "Experimental")

## 73 surveys 
## 34 control and 39 experimental 
## 41 group 1 control

group3$year <- as.character(group3$year)
group3$day <- as.character(group3$day)

summary <- full_join(group_1_2, group3) %>% filter(Calcium_mg_per_fisherman > 0)
## 232 rows (234 before filter fxn) 

summary %>% write.csv("alternate fishing/household_groups_finaldf.csv")

### Group 1 = 41 
### Group 2 = 118
### Group 3 = 73; 34 control and 39 experimental 
```

### Data exploration 

```{r}
summary2 <- summary %>%
  gather("measurement", "value", 7:25)

plot <- summary2 %>%
  ggplot(., aes(x=household_group, y=value, color=household_group)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=0.5, alpha=0.5, width=0.2) +
  facet_wrap(~measurement, scales = "free") + 
  theme_classic() + ylab("") + xlab("") +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x=element_blank())

ggsave(file="alternate fishing/metrics.png", plot, width = 17, height = 12, units = c("in"))

summary_stats <- summarySE(summary2, measurevar = c("value"), 
                           groupvars = c("household_group", "measurement"), na.rm = TRUE)

plot2 <- summary_stats %>%
  ggplot(., aes(x=household_group, y=value, color=household_group)) + 
  geom_errorbar(aes(x=household_group, y=value, ymin=value-se, ymax=value+se), 
                position=position_dodge(0.3), alpha=0.9, size=0.75, width=0.2) + 
  geom_point(size=3, position=position_dodge(0.3)) +
  facet_wrap(~factor(measurement, levels=c('Diet_percent_Carnivorous', 
                                           'Diet_percent_Herbivorous_Detritivorous', 
                                           'Diet_percent_Herbivorous_Macroalgal',
                                           'Diet_percent_Omnivorous',
                                           'Diet_percent_Planktivorous',
                                           'Iron_mg_per_fisherman',
                                           'Calcium_mg_per_fisherman',
                                           'Omega3_g_per_fisherman',
                                           'Protein_g_per_fisherman',
                                           'Selenium_ug_per_fisherman',
                                           'VitaminA_ug_per_fisherman',
                                           'Zinc_mg_per_fisherman',
                                           'Avg_Trophic_Level',
                                           'Mean_length_cm',
                                           'total_catch_per_fisherman',
                                           'richness',
                                           'Total_Biomass_kg_per_fisherman',
                                           'TakeHome_kg',
                                           'Total_value_KES_per_fisherman')), 
             labeller = as_labeller(c(Diet_percent_Carnivorous = "% of catch: Carnivorous",
                                      Diet_percent_Herbivorous_Detritivorous = "% of catch: Herbivorous Detritivorou",
                                      Diet_percent_Herbivorous_Macroalgal = "% of catch: Herbivorous Macroalgal",
                                      Diet_percent_Omnivorous = "% of catch: Omnivorous",
                                      Diet_percent_Planktivorous = "% of catch: Planktivorous",
                                      Iron_mg_per_fisherman = "Iron Content (mg) fisherman^-1",
                                      Calcium_mg_per_fisherman = "Calcium Content (mg) fisherman^-1",
                                      Omega3_g_per_fisherman = "Omega3 Content (g) fisherman^-1",
                                      Protein_g_per_fisherman = "Protein Content (g) fisherman^-1",
                                      Selenium_ug_per_fisherman = "Selenium Content (ug) fisherman^-1",
                                      VitaminA_ug_per_fisherman = "VitaminA Content (ug) fisherman^-1",
                                      Zinc_mg_per_fisherman = "Zinc Content (mg) fisherman^-1",
                                      Avg_Trophic_Level = "Trophic Level",
                                      Mean_length_cm = "Length (cm)",
                                      total_catch_per_fisherman = "Catch: # of fish fisherman^-1",
                                      richness = "# of unique species survey^-1",
                                      Total_Biomass_kg_per_fisherman = "Biomass (kg) fisherman^-1",
                                      TakeHome_kg = "Biomass (kg) taken home fisherman^-1",
                                      Total_value_KES_per_fisherman = "Value (KES) fisherman^-1")),
             scales = "free", strip.position = 'left') +
  ylab(NULL) + 
  xlab("") +
  theme(strip.background = element_blank(),
        strip.placement='outside',
        strip.text = element_text(face = "bold", size=10),
        axis.text.x=element_blank(),
        panel.background = element_rect(fill = 'white', color = 'black')) + 
  scale_color_manual(values = c("#53C936", "#E78818", "#48B5B7", "#185AE7")) +
  labs(color = "Household Group")

ggsave(file="alternate fishing/metrics2.png", plot2, width = 19, height = 12, units = c("in"))
```




