---
title: "Analysis of Fishing Landings dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal 

**1. Total catch per unit effort between modified and traditional traps. It would be great to see this as grams captured per trap set.**   

Results:    
- Modified traps had a significantly higher total catch per trap set. This pattern holds for all landing sites and individual fisherman (it does not appear that either of those variables influence this result).      
- Unmodified traps had a significantly higher weight in grams per trap set. This pattern holds for all individual fisherman, but Mayungu landing site might drive this result more than the other sites.      



2. Species catch per unit effort between modified and traditional traps. Take the top 3-5 species and run #1 for them separately.  
3. Total mature fish catch per unit effort between modified and traditional traps. This will have to be for the top 3-5 species separately. Go to Fishbase and find the length at first maturity for that particular species, then assign each fish a "mature" or "immature" status in the data and calculate.    
4. Average length of catch versus length at first maturity (Lmat). Take the difference for each fish in the data against its length at first maturity and then calculate a weighted value for modified versus traditional traps where a value above 0 represents a fish above Lmat and a value below represents a fish below Lmat.   
5. Length frequency of top 3-5 species in modified versus traditional (different colors) with Lmat etc. indicators pulled from Fishbase.

## Contents 

- [**Reading in datafiles**](#data)  
- [**Total catch (grams) per unit effort (trap set)**](#catch_effort)  

## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
library(stats)
```

Read in the data frame that is the output of the QC script. 

```{r, message=FALSE, warning=FALSE}
# read in excel file
data <- read_excel("data/Fishlandings-cleaned-21052022-May.xlsx") 
```

## <a name="catch_effort"></a> **Total catch (grams) per unit effort (trap set)**

Grouping by fisher_id but this might be effective to group by enumerator once I have correct list of names. There are 3 boat trips recorded with the exact same fish data that are under 3 different fisher ID names but all the same enumerator.. come back to this in QC.

Goal: grams captured per trap set.

```{r}
modified_trap_df <- data %>% unite(survey_id, Operation_date, fisher_id, sep = " ", remove = FALSE) %>%
  dplyr::group_by(survey_id) %>% 
  mutate(total_catch = sum(number_of_fish),
         grams_per_trap = total_weight_kg/total_traps_collected,
         catch_per_trap = total_catch/total_traps_collected) %>%
  dplyr::ungroup(survey_id) %>%
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>%
  select(survey_id, enumerator, trap_type, `No. of fishers in crew`, landing_site, total_catch, grams_per_trap, catch_per_trap) %>%
  distinct()
```

### Plotting figures.

Catch per trap 

```{r}
# basic total catch per trap with no other variables 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Total catch per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# does this differ by fisherman or landing site?
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~enumerator) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~landing_site) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle
```

Grams per trap 

```{r}
# basic grams per trap plot with no other variables 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# visually seeing if this differs by fisherman 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~enumerator) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# visually seeing if this differs by landing site 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~landing_site) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle
```

The relationship between grams per trap and total catch per trap. 

```{r}
modified_trap_df %>% 
  ggplot(aes(x=catch_per_trap, y=grams_per_trap, color=trap_type)) +
  theme_bw() + xlab("total catch per trap") + ylab("grams per trap") +
  geom_smooth(aes(fill=trap_type), method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
  geom_point(aes(fill=trap_type), pch = 21, size=1)
```

### Statistics on the above. 

#### Grams per trap and total catch 

Default of t.test in R is a Welch t-test which is just an adaptation of t-test, and it is used when the two samples have possibly unequal variances. Use var.equal = TRUE or FALSE to specifiy the variances of the dataset. 

You can test equal variances with a Fisher's F-test. If p < 0.05 then we include var.equal = FALSE in below ttest. If p > 0.05 then we include var.equal = TRUE in below ttest.

We are using an unpaired two sample t-test for this dataset. 

```{r}
UN <- modified_trap_df %>% subset(trap_type == "MODIFIED") %>% na.omit()
MOD <- modified_trap_df %>% subset(trap_type == "UNMODIFIED") %>% na.omit()
```

Grams per trap 

```{r}
var.test(UN$grams_per_trap, MOD$grams_per_trap)
t.test(grams_per_trap~trap_type, data = modified_trap_df, var.equal = FALSE)
```

Total catch per trap  

```{r}
var.test(UN$catch_per_trap, MOD$catch_per_trap)
t.test(catch_per_trap~trap_type, data = modified_trap_df, var.equal = FALSE)
```

Total catch per trap vs weight in grams per trap. 

```{r}

```


Left off at statistics and results for total catch per trap vs weight in grams per trap... question: does higher catch # result in higher weight in each modified and unmodified traps?

