---
title: "Analysis of Fishing Landings dataset - alternate dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Prior Scripts 

QC: https://github.com/emmastrand/Kenya_SamakiSalama/blob/main/FishLandings/scripts/QC.md 


## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
library(stats)
library(lme4)
library(car)
library(forcats)
library(ggstatsplot)
library(ggridges)
library(ggbreak)
library(emmeans)
```

## Read in the data frame that is the output of the QC script. 

Summary information so far: 741 total surveys, 591 unmodified traps, 150 modified traps. 

```{r, message=FALSE, warning=FALSE}
# read in excel file
## 2,156 rows x 30 variables (this should match the final version of QC script)
data <- read.csv("alternate fishing/alternate_fishing.csv") %>% #read in excel file 
  dplyr::select(-X)

# creating new columns with month year and date
## survey ID is master identifier that is based on `date_collected_dd_mm_yyyy` column 
data <- data %>% 
  separate(date_collected_dd_mm_yyyy, c("year", "month", "day"), remove = FALSE) 

#changing this column to numeric instead of a character (needed for next fxn)
data$month <- as.numeric(data$month) 
#changing numeric months to month names
data$month <- month.abb[data$month] 

# changing levels of month (important for figures later on); only first three letter of month 
data$month <- factor(data$month, levels=c("Jan", "Feb","Mar","Apr","May",
                                          "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data <- data %>%
  mutate(household_group = case_when(
    BMU == "KANAMAI" ~ "Control",
    BMU == "KURUWITU" ~ "Social marketing",
    BMU == "TAKAUNGU" ~ "Social marketing"))
```


**Control households**   
- KANAMAI (141 surveys): November 2021 and January - May 2022
- Gear used: Spearguns
- Units of data: Per single fisherman / household

**Social marketing households**   
- KURUWITU (382 surveys): July - December 2021 and January - May 2022  
- TAKAUNGU (105 surveys): January - May 2022  
- Gear used: Both used spearguns and TAKAUNGU used monofilament 
- Units of data: Per single fisherman / household

## Check ranges and add filters 

63 unique species total:  
- 40 of those found in KANAMAI  
- 56 of those found in KURUWITU  
- 32 of those found in TAKAUNGU  

```{r}
hist(data$total_biomass_kg) ## 0.25 26.00
hist(data$take_home_weight_kg, na.rm = TRUE) ## 0-4
hist(data$total_value_KES) ## 40 - 4800 
hist(data$take_home_value_KES, na.rm = TRUE) ## 0-600 

range(data$`No..of.fishers.in.crew`) ## 1-3 people 

hist(data$number_of_fish) ## 1 - 19 fish 

#unique(data$scientific_name) ## 63 fish 
#data %>% subset(BMU == "TAKAUNGU") %>% dplyr::select(scientific_name) %>% distinct()
```

## Calculating median length 

```{r}
data <- data %>% 
  mutate(median_length = case_when(
    length_corrected == "0-10" ~ 5,
    length_corrected == "11-15" ~ 13,
    length_corrected == "16-20" ~ 18,
    length_corrected == "21-25" ~ 23,
    length_corrected == "26-30" ~ 28,
    length_corrected == "31-35" ~ 33,
    length_corrected == "36-40" ~ 38,
    length_corrected == "41-45" ~ 43,
    length_corrected == "46-50" ~ 48,
    length_corrected == "51-60" ~ 55.5,
    length_corrected == "61-70" ~ 65.5,
    length_corrected == "71-80" ~ 75.5,
    length_corrected == "81-90" ~ 85.5,
    length_corrected == ">90" ~ 100 ### circle back to what value to use here
  )) 

hist(data$median_length) ## majority fall within 10-40 cm 
```


## Loading Galligan and FishBase dataset 

Most species in the list above have metadata below. 

```{r}
fishbase_lifehistory <- read_excel("data/fishbase.xlsx", sheet = "life history") %>% #read in excel file 
  select(scientific_name, Lm, Lmax) %>% dplyr::rename(Lmat_fishbase = Lm) %>% dplyr::rename(Lmax_fishbase = Lmax) 

# rows that appear in species_list but not fishbase_lifehistory
## checking which fish I still need Fishbase information for (44 species)
# dplyr::setdiff(species_list, fishbase_lifehistory) %>% write.csv("data/add_fishbase_info.csv")

fishbase_biomass <- read_excel("data/fishbase.xlsx", sheet = "biomass") %>%
  dplyr::select(scientific_name, a, b)

fishbase <- full_join(fishbase_lifehistory, fishbase_biomass, by = "scientific_name")

galligan <- read.csv("data/SpeciesData_GatedTraps_Galligan_edited.csv", header=TRUE, sep = ",") %>%
  dplyr::rename(scientific_name = Species) #%>% dplyr::select(., scientific_name)
# dplyr::setdiff(species_list, galligan) 
### 11 fish found in our dataset that are not found in Galligan dataset 
## these won't be included in the nutrient or revenue analysis
## I think this is OK b/c all species have 113 or lower than 40 total catch

metadata <- full_join(fishbase, galligan, by = "scientific_name")

### PRE MERGE DATA = 2,156 X 34
### POST MERGE DATA = 2,156 X 82 
data2 <- left_join(data, metadata, by = "scientific_name")
```

## Calculating total biomass 

W=aL^b 

W = Biomass  
L = Length (in our case, median length)  
a = (taken from fishbase)
b = (taken from fishbase)

Read in data from fishbase. These units are in cm and grams. 

https://www.fishbase.de/manual/English/FishbaseThe_LENGTH_WEIGHT_table.htm

```{r}
### Switching to data3 which will filter out those fish with no nutrient or a and b values 
data2 <- data2 %>%  
  ## mutating a and b column to be numeric; this will create NAs b/c not all fish had a and b valueus in fishbase
  mutate_at(c('a', 'b'), as.numeric) %>%
  mutate(W_g = (a*(median_length^b))*number_of_fish,
         W_kg = W_g/1000) %>%
  group_by(survey_id) %>%
  filter(!is.na(W_kg)) %>%
  mutate(calculated_total_biomass = sum(W_kg)) %>% 
  ungroup()

range(data2$calculated_total_biomass) ## 0.001832111 kg - 18.201910245 kg
```

### Comparing reported vs. calculated 

```{r}
data2 %>% 
  select(survey_id, total_biomass_kg, calculated_total_biomass) %>%
  distinct() %>% na.omit() %>%
  mutate(comparison = if_else(total_biomass_kg > calculated_total_biomass, "REPORTED > CALC", "REPORTED < CALC"),
         difference = total_biomass_kg-calculated_total_biomass) %>%
  ggplot(., aes(x=comparison, y=difference)) + theme_bw() +
  geom_jitter(aes(color=comparison), size=1) + xlab("") + ylab("difference in value")
```

### Statistics on the above. 

```{r}
data2_yield <- data2 %>% 
  dplyr::select(survey_id, household_group, total_biomass_kg, calculated_total_biomass) %>% distinct()

Control_yield <- data2_yield %>% subset(household_group == "Control") ## 131 data points
Social_yield <- data2_yield %>% subset(household_group == "Social marketing") ## 448 data points 

### TOTAL BIOMASS KG 
var.test(Control_yield$total_biomass_kg, Social_yield$total_biomass_kg) 
## p-value = 5.004e-06; ratio = 0.5002933
t.test(total_biomass_kg~household_group, data = data2_yield, var.equal = FALSE)
## p-value = 0.001342

### TOTAL CALCULATED BIOMASS 
var.test(Control_yield$calculated_total_biomass, Social_yield$calculated_total_biomass) 
## p-value = 0.0323; ratio = 0.7299033
t.test(calculated_total_biomass~household_group, data = data2_yield, var.equal = FALSE)
## p-value = 9.33e-09
```

### Figures 

```{r}
calculated_yield_stats <- summarySE(data2_yield, measurevar = c("calculated_total_biomass"), groupvars = c("household_group"))
total_biomass_stats2 <- summarySE(data2_yield, measurevar = c("total_biomass_kg"), groupvars = c("household_group"))

calculated_yield_stats %>%
  ggplot(., aes(x=household_group, y=calculated_total_biomass, fill=household_group, group=household_group)) + theme_bw() +
  geom_errorbar(aes(x=household_group, y=calculated_total_biomass, 
                    ymin=calculated_total_biomass-se, ymax=calculated_total_biomass+se), 
                position=position_dodge(0.3), alpha=0.9, size=0.5, width=0.1) +  
  geom_point(size=6, shape=21, position=position_dodge(0.3)) + ylab("Calculated Total Biomass") +
  xlab("Household Group") + labs(fill="Household Group")
# 
# total_biomass_stats2 %>%
#   ggplot(., aes(x=household_group, y=total_biomass_kg, fill=household_group, group=household_group)) + theme_bw() +
#   geom_errorbar(aes(x=household_group, y=total_biomass_kg, 
#                     ymin=total_biomass_kg-se, ymax=total_biomass_kg+se), 
#                 position=position_dodge(0.3), alpha=0.9, size=0.5, width=0.1) +  
#   #scale_y_break(c(1.1, 1.75)) + ylim(0.9,2.4) +
#   geom_point(size=3, shape=21, position=position_dodge(0.3)) 
```

## Calculating take home kg/% 

```{r}
takehome <- data2 %>% dplyr::group_by(survey_id, destination) %>%   
  mutate(kg_perdestination = sum(W_kg)) %>% ungroup() %>% 
  dplyr::select(survey_id, BMU, household_group, destination, kg_perdestination) %>%
  distinct() %>% spread(destination, kg_perdestination) %>%
  mutate_all(., ~replace_na(.,0)) %>% dplyr::select(survey_id, BMU, household_group, HOME) %>%
  dplyr::rename(TakeHome_kg = HOME)
```

### Statistics on the above. 

```{r}
data2_takehome <- takehome %>% 
  dplyr::select(survey_id, household_group, TakeHome_kg) %>% distinct() %>% 
  filter(TakeHome_kg > 0) ##99 total observations #23 control #76 marketing

Control_takehome <- data2_takehome %>% subset(household_group == "Control") ## 131 total data points
Social_takehome <- data2_takehome %>% subset(household_group == "Social marketing") ## 448 total data points 

### TOTAL BIOMASS KG 
var.test(Control_takehome$TakeHome_kg, Social_takehome$TakeHome_kg) 
## p-value = 8.383e-05; ratio = 0.198125
t.test(TakeHome_kg~household_group, data = data2_takehome, var.equal = FALSE)
## p-value = 0.2809
```

### Figures 

```{r}
#data2_takehome_nozero <- data2_takehome %>% filter(TakeHome_kg > 0) ##99 total observations #23 control #76 marketing

TakeHome_stats <- summarySE(data2_takehome, measurevar = c("TakeHome_kg"), groupvars = c("household_group"), 
                           na.rm = TRUE)
#TakeHome_stats %>%
data2_takehome %>%
  ggplot(., aes(x=household_group, y=TakeHome_kg, fill=household_group, group=household_group)) + theme_bw() +
  # geom_errorbar(aes(x=household_group, y=TakeHome_kg,
  #                   ymin=TakeHome_kg-se, ymax=TakeHome_kg+se),
  #               position=position_dodge(0.3), alpha=0.9, size=0.5, width=0.1) +
  #geom_point(size=6, shape=21, position=position_dodge(0.3)) + 
  geom_boxplot() + 
  geom_jitter(size=1, shape=21, width = 0.2) + 
  ylab("Take Home Biomass (kg)") +
  xlab("Household Group") + labs(fill="Household Group")
```

## Calculating trophic level 

```{r}
data2 <- data2 %>% group_by(survey_id) %>%
  mutate(mean.trophic = mean(TrophLevel, na.rm = TRUE))
```

### Statistics on the above. 

```{r}
data2_trophic <- data2 %>% 
  dplyr::select(survey_id, household_group, mean.trophic) %>% distinct()

Control_trophic <- data2_trophic %>% subset(household_group == "Control") ## 131 data points
Social_trophic <- data2_trophic %>% subset(household_group == "Social marketing") ## 448 data points 

### TOTAL BIOMASS KG 
var.test(Control_trophic$mean.trophic, Social_trophic$mean.trophic) 
## p-value = 0.3708; ratio = 1.129964
t.test(mean.trophic~household_group, data = data2_trophic, var.equal = TRUE)
## p-value = 0.1518
```

### Figures 

```{r}
trophic_stats <- summarySE(data2_trophic, measurevar = c("mean.trophic"), groupvars = c("household_group"), 
                           na.rm = TRUE)

trophic_stats %>%
  ggplot(., aes(x=household_group, y=mean.trophic, fill=household_group, group=household_group)) + theme_bw() +
  geom_errorbar(aes(x=household_group, y=mean.trophic, 
                    ymin=mean.trophic-se, ymax=mean.trophic+se), 
                position=position_dodge(0.3), alpha=0.9, size=0.5, width=0.1) +  
  geom_point(size=6, shape=21, position=position_dodge(0.3)) + 
  ylab("Mean Trophic Level of Catch") +
  xlab("Household Group") + labs(fill="Household Group")
```


## Calculating herbivore/carnivore presence 

```{r}
diet <- data2 %>% group_by(survey_id, Diet) %>%
  mutate(countper_diet = n()) %>% ungroup() %>%
  select(survey_id, BMU, household_group, Diet, countper_diet) %>% distinct() %>%
  group_by(survey_id) %>% spread(Diet, countper_diet) %>% dplyr::select(-`<NA>`) %>%
  mutate_all(., ~replace_na(.,0)) %>%
  mutate(total = sum(c(Carnivorous, `Herbivorous detritivorous`, `Macroalgal herbivorous`,
                        `Mobile Inverts`, `Omnivorous`, `Planktivorous`))) %>%
  mutate(Carnivorous = Carnivorous/total*100,
         `Herbivorous detritivorous` = `Herbivorous detritivorous`/total*100,
         `Macroalgal herbivorous` = `Macroalgal herbivorous`/total*100,
         `Mobile Inverts` = `Mobile Inverts`/total*100,
         `Omnivorous` = `Omnivorous`/total*100,
         `Planktivorous` = `Planktivorous`/total*100) %>% dplyr::select(-total)
  
head(diet)
```

## Summary 

```{r}
summary <- data2 %>% 
  dplyr::select(survey_id, BMU, household_group, calculated_total_biomass, mean.trophic) %>%
  distinct() %>% dplyr::rename(Total_Biomass_kg = calculated_total_biomass) %>%
  dplyr::rename(Avg_Trophic_Level = mean.trophic) %>%
  left_join(., takehome, c("survey_id", "BMU", "household_group")) %>%
  left_join(., diet, by = c("survey_id", "BMU", "household_group")) 

head(summary)
```





