---
title: "CUE, maturity, and length analysis of Fishing Landings dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions for Austin and Clay 

## Summary of dataset 

Modified traps data:  
2022: January, February, March, April, May  

Unmodified traps data:
2021: January, June, August, September
2022: January, February, March, April, May, June

## Contents 

- [**Reading in datafiles**](#data)  
- [**Total catch (grams) per unit effort (trap set)**](#catch_effort)  
- [**Calculate top species caught**](#species)  
- [**Top species stats per trap**](#species_pertrap)  
- [**Creating database from Fishbase**](#fishbase)  
- [**Catch per unit effort for top species by maturity**](#maturity)  
- [**Catch and length data of mature fish**](#length)   
- [**Length Frequency plots of top species**](#freq_plots)   

## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
library(stats)
library(lme4)
library(car)
```

Read in the data frame that is the output of the QC script. 

```{r, message=FALSE, warning=FALSE}
# read in excel file
data <- read_excel("data/Fishlandings-cleaned-clay-June_updated-IW.xlsx") #read in excel file 

data <- data %>% separate(Operation_date, c("year", "month", "day"), remove = FALSE) # creating new columns with month year and date
data$month <- as.numeric(data$month) #changing this column to numeric instead of a character (needed for next fxn)
data$month <- month.abb[data$month] #changing numeric months to month names 
```

## <a name="catch_effort"></a> **Total catch (grams) per unit effort (trap set)**

**Total catch per unit effort between modified and traditional traps. It would be great to see this as grams captured per trap set.**

Grouping by fisher_id but this might be effective to group by enumerator once I have correct list of names. There are 3 boat trips recorded with the exact same fish data that are under 3 different fisher ID names but all the same enumerator.. come back to this in QC.

Goal: grams captured per trap set.

```{r}
modified_trap_df <- data %>% unite(survey_id, Operation_date, fisher_id, sep = " ", remove = FALSE) %>%
  dplyr::group_by(survey_id) %>% # group by survey id
  mutate(total_catch = sum(number_of_fish), #count the number of fish caught for each survey id
         grams_per_trap = total_weight_kg/total_traps_collected, #divide total weight for survey id by total traps 
         catch_per_trap = total_catch/total_traps_collected) %>% #divide total catch per survey id by total traps 
  dplyr::ungroup(survey_id) %>% #ungroup by this column  
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>% #subset for only modified and unmodified traps 
  select(survey_id, enumerator, trap_type, `No. of fishers in crew`, landing_site, total_catch, month, year, grams_per_trap, catch_per_trap) %>%
  distinct() 
```

### Plotting figures.

Catch per trap 

```{r}
# basic total catch per trap with no other variables 
modified_trap_df %>% filter(catch_per_trap < 1500) %>%
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Total catch per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by fisherman 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~enumerator) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Total catch per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by landing site
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~landing_site) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Total catch per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by month and year 
modified_trap_df %>% unite(ym, year, month, sep = " ", remove = FALSE) %>% filter(catch_per_trap < 1500) %>%
  ggplot(aes(x=month, y=catch_per_trap, color=trap_type)) + 
  facet_wrap(~year, scales = "free_y") +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
  #geom_point(aes(x=ym, y=catch_per_trap, fill=trap_type), pch = 21, size=1) +
  theme_classic() + 
  ylab("Total catch per trap") + xlab("Time of year") +
  theme(axis.text.x = element_text(vjust = 1.1, hjust=1.1, angle=60)) #Set the text angle
```

Grams per trap 

```{r}
# basic grams per trap plot with no other variables 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# visually seeing if this differs by fisherman 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~enumerator) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# visually seeing if this differs by landing site 
modified_trap_df %>% 
  ggplot(aes(x=trap_type, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~landing_site) +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
    geom_point(aes(fill=trap_type), pch = 21, size=1) +
  theme_bw() + 
  ylab("Grams of fish per trap") + xlab("Type of trap") +
  theme(axis.text.x = element_text(vjust = 1.1)) #Set the text angle

# by month and year 
modified_trap_df %>% unite(ym, year, month, sep = " ", remove = FALSE) %>% #filter(catch_per_trap < 1500) %>%
  ggplot(aes(x=month, y=grams_per_trap, color=trap_type)) + 
  facet_wrap(~year, scales = "free_y") +
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
 # geom_point(aes(x=month, group=trap_type, y=grams_per_trap, fill=trap_type), pch = 21, size=1) +
  theme_classic() + 
  ylab("Grams of fish per trap") + xlab("Time of year") +
  theme(axis.text.x = element_text(vjust = 1.1, hjust=1.1, angle=60)) #Set the text angle
```

The relationship between grams per trap and total catch per trap. 

```{r}
modified_trap_df %>% filter(catch_per_trap < 1500) %>%
  ggplot(aes(x=catch_per_trap, y=grams_per_trap, color=trap_type)) +
  theme_bw() + xlab("total catch per trap") + ylab("grams per trap") +
  geom_smooth(aes(fill=trap_type), method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
  geom_point(aes(fill=trap_type), pch = 21, size=1)
```

### Statistics on the above. 

Default of t.test in R is a Welch t-test which is just an adaptation of t-test, and it is used when the two samples have possibly unequal variances. Use var.equal = TRUE or FALSE to specifiy the variances of the dataset. 

You can test equal variances with a Fisher's F-test. If p < 0.05 then we include var.equal = FALSE in below ttest. If p > 0.05 then we include var.equal = TRUE in below ttest.

We are using an unpaired two sample t-test for this dataset. 

```{r}
UN <- modified_trap_df %>% subset(trap_type == "MODIFIED") %>% filter(!is.na(grams_per_trap)) %>% filter(!is.na(catch_per_trap))
MOD <- modified_trap_df %>% subset(trap_type == "UNMODIFIED") %>% filter(!is.na(grams_per_trap)) %>% filter(!is.na(catch_per_trap))
```

#### Grams per trap 

```{r}
var.test(UN$grams_per_trap, MOD$grams_per_trap)
t.test(grams_per_trap~trap_type, data = modified_trap_df, var.equal = FALSE)
```

#### Total catch per trap  

```{r}
var.test(UN$catch_per_trap, MOD$catch_per_trap)
t.test(catch_per_trap~trap_type, data = modified_trap_df, var.equal = FALSE)
```

#### Total catch per trap vs weight in grams per trap. 

We use a linear mixed model for this so we can include other variables like fisherman and landing site. 

Grams per trap is log transformed.

I'm not sure yet if this is the best way to do this... 

```{r}
# unmodified
un_catch_model <- lmer(log(grams_per_trap) ~ catch_per_trap + (1|enumerator) + (1|landing_site), na.action=na.omit, data=UN)
qqPlot(residuals(un_catch_model)) 
hist(residuals(un_catch_model))

# modified
mod_catch_model <- lmer(log(grams_per_trap) ~ catch_per_trap + (1|enumerator) + (1|landing_site), na.action=na.omit, data=MOD)
qqPlot(residuals(mod_catch_model)) 
hist(residuals(mod_catch_model))

summary(un_catch_model)
summary(mod_catch_model)

Anova(un_catch_model, ddf="lme4", type='III')
Anova(mod_catch_model, ddf="lme4", type='III')
```

## <a name="species"></a> **Calculate top species caught**

**Species catch per unit effort between modified and traditional traps. Take the top 3-5 species and run #1 for them separately.**

Calculating which species were the most abundant across the entire survey.  

This might have to be number of fish caught per trap? So that the difference in # of traps for modified and unmodified is accounted for?

This is split for modified and unmodified so far.. but can be changed to combined.. the trend is about the same for most abundant type of fish in each trap.. 

```{r}
species_list <- data %>% select(scientific_name, number_of_fish, trap_type) %>% 
  filter(!is.na(trap_type)) %>% 
  filter(!is.na(number_of_fish)) %>% 
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>% #subset for only modified and unmodified traps 
  dplyr::group_by(scientific_name, trap_type) %>%
  mutate(species_catch = sum(number_of_fish)) %>% 
  select(-number_of_fish) %>% distinct() %>%
  ungroup()

# above 250 cut off includes top each count for each type of trap 
species_list %>% filter(species_catch > 250) %>%
  ggplot(., aes(x=scientific_name, y=species_catch, group = trap_type, color = trap_type)) + 
  ylab("number of fish caught") + xlab("Genus species") +
  geom_point() + theme_bw() + theme(axis.text.x = element_text(angle = 60, hjust=1)) #Set the text angle

# species_list %>% filter(species_catch > 250) %>% 
#   select(scientific_name) %>% 
#   distinct() %>%
#   write_csv("output/top_species.csv")

# print top 5 species from modified traps 
species_list %>% subset(trap_type == "MODIFIED") %>%                                    
  arrange(desc(species_catch)) %>% head(5)

# print top 5 species from unmodified traps 
species_list %>% subset(trap_type == "UNMODIFIED") %>%                                    
  arrange(desc(species_catch)) %>% head(5)
```

## <a name="species_pertrap"></a> **Top species stats per trap**

Create a subsetted df from the top 10 total species (break this down into modified and unmodified later?).

```{r}
# calculate species total catch to then figure out which ones were most popular regardless of trap type
species_df <- data %>% filter(!is.na(number_of_fish)) %>%
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>%
  group_by(scientific_name) %>%
  mutate(species_total_catch = sum(number_of_fish)) %>% ungroup() #must ungroup for following commands 

# use the above metric to subset to top 10 of those 
species_keep <- species_df %>% select(scientific_name, species_total_catch) %>% 
  distinct() %>% slice_max(species_total_catch, n = 10) 

# filter species df based on the species_keep list 
species_df <- species_df %>% filter(scientific_name %in% species_keep$scientific_name)

# double checking the above command worked - output should be only 5 
unique(sort(species_df$scientific_name))
```

With just the top 10 species, just plotting catch per trap. I don't think I can do weight of just these species by trap because the weight is for the whole catch. 

```{r}
### species_df is already subsetted to the top 10 species 
# within each survey id, for each species calculate topspecies_catch as the number fish in that species in that survey
# then calculate the catch per trap as the number of fish of that species over the total number of traps collected
species_df <- species_df %>% unite(survey_id, Operation_date, fisher_id, sep = " ", remove = FALSE) %>%
  dplyr::group_by(survey_id, scientific_name) %>% # group by survey id
  mutate(topspecies_catch = sum(number_of_fish),
         catch_per_trap = topspecies_catch/total_traps_collected) %>% #divide total catch per survey id by total traps 
  dplyr::ungroup(survey_id, scientific_name) #ungroup by this column

# basic grams per trap plot with no other variables 
species_df %>% mutate(trap_type = case_when(
    trap_type == "MODIFIED" ~ "MOD",
    trap_type == "UNMODIFIED" ~ "UNMOD")) %>%
  ggplot(aes(x=trap_type, y=catch_per_trap, color=trap_type)) + 
  geom_boxplot(aes(color=trap_type), outlier.size = 0, lwd=0.5) +
  facet_wrap(~scientific_name, scales = "free_y") +
  theme_classic() + ggtitle("For top 10 most abundant species") +
  ylab("Catch per trap (species catch # / total catch)") + xlab("Genus species") +
  theme(axis.text.x = element_text()) #Set the text angle
```

### Statistics for the top ten species

Left off at trying to solve the following error when running aov: 

Error in as.POSIXlt.character(x, tz, ...) : character string is not in a standard unambiguous format

Error in lm.fit(x, y, offset = offset, singular.ok = singular.ok, ...) : NA/NaN/Inf in 'y'

```{r}
Topspp_MOD <- species_df %>% subset(trap_type == "MODIFIED")

Topspp_UNMOD <- species_df %>% subset(trap_type == "UNMODIFIED")  

var.test(Topspp_MOD$catch_per_trap, MOD$catch_per_trap) # output is significant so put false below

#t.test(catch_per_trap~trap_type, data = species_df, var.equal = FALSE)

#aov(catch_per_trap ~ trap_type + scientific_name, data = species_df)
```

## <a name="maturity"></a> **Catch per unit effort for top species by maturity**

**Total mature fish catch per unit effort between modified and traditional traps. This will have to be for the top 3-5 species separately. Go to Fishbase and find the length at first maturity for that particular species, then assign each fish a "mature" or "immature" status in the data and calculate.**

Data pulled from Fishbase and put in the datasheet `fishbase.xlsx`. 

### Creating maturity dataframe based on data pulled from fishbase

```{r}
fishbase <- read_excel("data/fishbase.xlsx", sheet = "data") %>% #read in excel file 
  select(scientific_name, Lm, Lm_se_min, Lm_se_max)

maturity <- full_join(data, fishbase, by = "scientific_name") %>% # joining maturity reproduction info with main dataframe
  unite(survey_id, Operation_date, fisher_id, sep = " ", remove = FALSE) %>% #creating a survey id based on date and fisherman
  filter(!is.na(Lm)) %>% # taking out observations that don't have an Lm value so only the top species 
  mutate(Lm_range = case_when( ### making a new column as Lm corrected to then compare to Length_corrected 
    Lm >= 0 & Lm <= 10.5 ~ "0-10",
    Lm >= 10.5 & Lm <= 15.4 ~ "11-15",
    Lm >= 15.5 & Lm <= 20.4 ~ "16-20",
    Lm >= 20.5 & Lm <= 25.4 ~ "21-25",
    Lm >= 25.5 & Lm <= 30.4 ~ "26-30",
    Lm >= 30.5 & Lm <= 35.4 ~ "31-35",
    Lm >= 35.5 & Lm <= 40.4 ~ "36-40",
    Lm >= 40.5 & Lm <= 45.4 ~ "41-45",
    Lm >= 45.5 & Lm <= 50.4 ~ "46-50",
    Lm >= 50.5 & Lm <= 75 ~ ">50",
    Lm > 75 ~ ">75")) %>%
  mutate(Lm_comparison = case_when( ### I can't compare character columns so making a new column to use in comparison
    Lm_range == "0-10" ~ 1,
    Lm_range == "11-15" ~ 11,
    Lm_range == "16-20" ~ 16,
    Lm_range == "21-25" ~ 21,
    Lm_range == "26-30" ~ 26,
    Lm_range == "31-35" ~ 31,
    Lm_range == "36-40" ~ 36,
    Lm_range == "41-45" ~ 41,
    Lm_range == "46-50" ~ 46,
    Lm_range == ">50" ~ 50,
    Lm_range == ">75" ~ 75)) %>%
  mutate(Length_comparison = case_when( ### I can't compare character columns so making a new column to use in comparison
    length_corrected == "0-10" ~ 1,
    length_corrected == "11-15" ~ 11,
    length_corrected == "16-20" ~ 16,
    length_corrected == "21-25" ~ 21,
    length_corrected == "26-30" ~ 26,
    length_corrected == "31-35" ~ 31,
    length_corrected == "36-40" ~ 36,
    length_corrected == "41-45" ~ 41,
    length_corrected == "46-50" ~ 46,
    length_corrected == ">50" ~ 50,
    length_corrected == ">75" ~ 75))
  
### there is probably a more efficient way to do the above this works for now 

maturity$Length_comparison <- as.numeric(maturity$Length_comparison)
maturity$Lm_comparison <- as.numeric(maturity$Lm_comparison)

maturity <- maturity %>% 
  mutate(maturity = if_else(length_corrected >= Lm_comparison, "mature", "immature")) %>%
  subset(trap_type == "MODIFIED" | trap_type == "UNMODIFIED") %>% #subsetting to only the modified and unmodified traps 
  mutate(nofish_pertrap = number_of_fish/total_traps_collected) #per species per survey id
```

Immaure and mature fish catch comparison

```{r}
maturity %>% select(number_of_fish, trap_type, maturity, scientific_name, total_traps_collected, nofish_pertrap) %>% 
  na.omit() %>% 
  ggplot(., aes(x=scientific_name, y=nofish_pertrap, color=maturity)) +
  geom_boxplot() + theme_classic() + 
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  xlab("Species") +
  ylab("Number of fish per trap set")

maturity %>% select(number_of_fish, trap_type, maturity, scientific_name, total_traps_collected, nofish_pertrap) %>% 
  na.omit() %>%  
  ggplot(., aes(x=maturity, y=nofish_pertrap, color=trap_type)) +
  geom_boxplot() + theme_classic() + 
  facet_wrap(~trap_type, scales = "free_y") +
  #theme(axis.text.x = element_text(angle=60, hjust=1)) +
  xlab("Maturity") +
  ylab("Number of fish per trap set")
```

Statistics on the above... 

Circle back to the padj values coming out to "0.000000e+00". 

Double check `trap_type+maturity` vs `trap_type*maturity`.  

```{r}
nofish_maturity_aov <- aov(number_of_fish ~ trap_type*maturity, data = maturity)
summary(nofish_maturity_aov)

nofish_maturity_tukey <- TukeyHSD(nofish_maturity_aov)
nofish_maturity_tukey$`trap_type:maturity`
```



## <a name="length"></a> **Catch and length data of mature fish**

**Average length of catch versus length at first maturity (Lmat). Take the difference for each fish in the data against its length at first maturity and then calculate a weighted value for modified versus traditional traps where a value above 0 represents a fish above Lmat and a value below represents a fish below Lmat.**

We only have bins for the length values.. so this might have to be by 5s. 
i.e. catch = 11-15; Lm 21-25. Can take median values and do that calculation? 

I'm not sure this is the best way to do this... 

Transform count value to that number of observations: https://stackoverflow.com/questions/70759069/transform-count-column-into-number-of-rows. 

```{r}
maturity_dist <- maturity %>% 
  select(survey_id, number_of_fish, scientific_name, trap_type, length_corrected, Lm, Lm_range, Lm_comparison, Length_comparison, maturity) %>%
  mutate(number_of_fish = if_else(is.na(number_of_fish), 1, number_of_fish)) %>% # replacing NAs with the value of 1
  tidyr::uncount(., number_of_fish, .remove = TRUE) %>% # expanding number of fish to number of observations
  #### notes on below fxn --- maybe not the best way to do this? 0-10 is a bigger bin than 11-15 so median value isn't as representative?
  mutate(median_length = case_when(
    length_corrected == "0-10" ~ 5,
    length_corrected == "11-15" ~ 13,
    length_corrected == "16-20" ~ 18,
    length_corrected == "21-25" ~ 23,
    length_corrected == "26-30" ~ 28,
    length_corrected == "31-35" ~ 33,
    length_corrected == "36-40" ~ 38,
    length_corrected == "41-45" ~ 43,
    length_corrected == "46-50" ~ 48,
    length_corrected == ">50" ~ 50,
    length_corrected == ">75" ~ 75)) %>%
  mutate(length_dist = median_length-Lm) %>%
  group_by(scientific_name, trap_type) %>%
  mutate(avg_length_dist = mean(length_dist)) %>% #### a positive value here = mature; a negative value = immature 
  ungroup()


maturity_dist %>% group_by(scientific_name) %>%
  filter(., n_distinct(avg_length_dist) >= 2) %>% # filters out the observations aren't in both categories (takes out species that only have 1 unique mean dist value so only 1 trap type)
  ungroup() %>%
  ggplot(., aes(x=scientific_name, y=avg_length_dist, color = trap_type)) +
  geom_line(aes(group = scientific_name), color="grey14") +
  geom_point() + theme_bw() + 
  geom_hline(yintercept = 0, lty = "dotted") +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  labs(y="Mean distance from Lm", x="Genus species", color="Trap Type")
```



## <a name="freq_plots"></a> **Length Frequency plots of top species**

**Length frequency of top 3-5 species in modified versus traditional (different colors) with Lmat etc. indicators pulled from Fishbase.**


Check this calculation for each fish not just each survey id???


http://derekogle.com/fishR/2017-07-28-JoyPlot

From the `species_keep` df:

Siganus sutor	= 299920  			
Lethrinus nebulosus	= 27161	  		
Scarus ghobban	= 23025			  
Leptoscarus vaigiensis	= 14070		  	
Siganus canaliculatus	= 9509	  

```{r}
# the column we want to plot: maturity$length_corrected
maturity$length_corrected <- factor(maturity$length_corrected, levels=c("0-10", "11-15","16-20","21-25","26-30","31-35","36-40",
                                                                        "41-45", "46-50", ">75"))

maturity <- maturity %>% filter(!is.na(length_corrected)) %>%
  group_by(length_corrected, scientific_name, trap_type) %>%
  mutate(count.per.bin = sum(number_of_fish)) %>%
  ungroup()


maturity2022_topspp <- maturity %>% subset(year=="2022") %>% 
  subset(scientific_name == "Siganus sutor" | scientific_name == "Lethrinus nebulosus" |
           scientific_name == "Scarus ghobban" | scientific_name == "Leptoscarus vaigiensis" |
           scientific_name == "Siganus canaliculatus")
  
maturity2022_topspp %>% 
  mutate(number_of_fish = if_else(is.na(number_of_fish), 1, number_of_fish)) %>% # replacing NAs with the value of 1
  tidyr::uncount(., number_of_fish, .remove = TRUE) %>% # expanding number of fish to number of observations
  ggplot(., aes(x=length_corrected, fill=trap_type, color=trap_type)) + 
  geom_bar(alpha=0.3) +
  geom_vline(data = maturity2022_topspp, mapping = aes(xintercept = Lm_range), lty = "dotted") +
  theme_bw() + xlab("Length (cm)") + ylab("Frequency") +
  theme(axis.text.x.bottom = element_text(colour = 'black', angle = 60, hjust = 1),
        axis.text.y = element_text(colour = 'black', size = 8, face = 'italic')) + 
  theme(panel.border = element_blank(),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   strip.text.x = element_text(size = 9, color = "black", face = "bold.italic"),
   axis.line = element_line(colour = "black")) +
  theme(axis.title.y = element_text(size = 12, face = "bold", margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  facet_wrap(~scientific_name, scales = "free_y") +
  ggtitle("Siganus canaliculatus Lm range = 0-10, can edit on in other program")
```










