---
title: "Quality Control of Fishing Landings dataset"
author: "Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to run this with a future xlsx file

1. Replace raw data file name in the function that creates the following variables: fishing_operation, catch_composition, and validation_lists.  

## Load all libraries 

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
```

## Create dataframe 

**Load in raw data files.** 

[Reading in an excel datafile instead of a csv](http://www.sthda.com/english/wiki/reading-data-from-excel-files-xls-xlsx-into-r). This requires the `readxl` package listed in load all libraries step. 

```{r, message=FALSE, warning=FALSE}
fishing_operation <- read_excel("data/Fishlandings-data_21052022-May_update-cleaning in progress.xlsx", sheet = "fishing_operation") %>%
  select(-'...21', -'...22', -'...23') %>%
  rename(Operation_date = date_dd_mm_yyyy)

catch_composition <- read_excel("data/Fishlandings-data_21052022-May_update-cleaning in progress.xlsx", sheet = "catch_composition") %>%
  select(-'...10', -'...11', -'...12', -'...13') %>%
  rename(Operation_date = Date)

validation_lists <- read_excel("data/Fishlandings-data_21052022-May_update-cleaning in progress.xlsx", sheet = "validation_lists")

df <- full_join(fishing_operation, catch_composition, by = c("fisher_id", "Operation_date")) 
```

## Quality Control 

```{r}
head(df)
```


**Operation_date**: in the above chunk, these dates are automatically read in as *dttm* format (date and time). 

**enumerator** 

```{r}

```



