---
title: "Relative abundance analysis of Fishing Landings dataset"
author: "Author: Emma Strand; emma_strand@uri.edu"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aims and Results 

**1. Quantify and visualize relative abundance for each modified and unmodified trap using phyloseq.** 

**2. Alpha and beta diversity for fish species between modified and unmodified traps. Does using modified traps affect the diversity of fish caught?**  

## Contents 

- [**Reading in datafiles**](#data)  

## <a name="data"></a> **Reading in datafiles**

```{r, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(Hmisc)
library(writexl)
library(naniar)
library(Rmisc)
library(stats)
library(lme4)
library(car)
```

Read in the data frame that is the output of the QC script and create matrix for phyloseq input. 

```{r}

```

